---
title: "Notepad-Examples"
author: "`r Sys.getenv('USERNAME')`"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
## basic settings
knitr::opts_chunk$set(echo = FALSE,
                      message=FALSE,
                      warning=FALSE
                      )

## load common libraries
library(tidyverse)
library(lubridate)
library(scales)
library(plotly)
library(PerformanceAnalytics)

## set variables
color.bar <- "cornflowerblue"

## Reference: R Markdown cheatsheet
## https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf


```

## Get some data

```{r LOAD DATA}
## grab data -> can use vancouver weather data in other R project
# data <- read_csv(file.choose())
# write_csv(data, 'data-input/vw-data.csv')

## set up to use data
idata <- read_csv('data-input/vw-data.csv')

## filter data for investigation
idata2 <- idata %>% filter(Year>=2017, Month==4) ## Daily April data for recent years

```


## Explore summary

```{r SUMMARY 1}

## summary stats for all
summary(idata2)

## summary statistics for each month
idata2.mth <- idata2 %>% group_by(Year, Month) %>% summarize(
  max.mean=mean(Max.Temp),
  min.mean=mean(Min.Temp),
  mean.mean=mean(Mean.Temp),
  mean.precip=mean(Total.Precip)
) 
idata2.mth

chart.title <- "Mean temperatures for same month, each year"
ggplot(idata2.mth, aes(x=Year, y=mean.mean))+geom_col(fill=color.bar)+
  scale_y_continuous(expand=c(0,0))+
  theme_minimal()+
  labs(title=chart.title)

```

### Distribution: what is the distribution for the temperatures in each month?

```{r DISTRIB 1}
## calculate mean of mean temp and std dev
idata2.mth.mt <- idata2 %>% group_by(Year) %>% summarize(
  mean.meant=mean(Mean.Temp),
  sd.meant=sd(Mean.Temp)
)
idata2.mth.mt <- idata2.mth.mt %>% mutate(
  onesd.below=mean.meant-sd.meant,
  onesd.above=mean.meant+sd.meant
)
idata2.mth.mt

```

```{r DISTRIB VIZ 1}
chart.title <- "Mean temperatures: mean +/- 1 st dev"
ggplot(idata2.mth.mt, aes(x=Year, y=mean.meant))+geom_col(fill=color.bar)+
  geom_errorbar(ymin=idata2.mth.mt$onesd.below, 
                                   ymax=idata2.mth.mt$onesd.above, width=0.25)+
   scale_y_continuous(expand=c(0,0), limits=c(0,max(idata2.mth.mt$onesd.above)+0.5))+
  theme_minimal()+
  labs(title=chart.title)

chart.title <- "Mean temp distribution each yr"
ggplot(idata2, aes(x=as.factor(Year), y=Mean.Temp))+geom_boxplot(fill=color.bar)+
  theme_minimal()+
  labs(title=chart.title)

```

### Are the mean temperatures different at a level of statistical significance?

Easy answer: 'Yes', because we have full 'population' data and so we have the 'true' means to compare.

The idea here is to apply stastistical methods as if it were sampled data. 

#### Histograms

##### Faceted

Dotted lines = mean temp

```{r HIST1}

ggplot(idata2, aes(x=Mean.Temp))+geom_histogram(fill=color.bar)+
  geom_vline(data=idata2.mth.mt, aes(xintercept=mean.meant), linetype='dashed')+
  facet_grid(Year~.)+theme_bw()

```

##### Overlaid

```{r HIST2}
ggplot(idata2, aes(x=Mean.Temp, fill=as.factor(Year)))+geom_histogram(alpha=0.5)+
  theme_minimal()
```

#### Density plot comparison

```{r DENSITY}
ggplot(idata2, aes(x=Mean.Temp, color=as.factor(Year)))+geom_density(size=1)+
  theme_minimal()
```

#### Statistical Testing

* Create separate datasets with mean temp for each period
* Apply t-test to each dataset individually first to test hypotheses about the mean

```{r TTEST DATA, echo=TRUE}
## separate out years
idata2.1 <- idata2 %>% filter(Year==2017)
idata2.2 <- idata2 %>% filter(Year==2018)
idata2.3 <- idata2 %>% filter(Year==2019)
```

T-test: sample 1

```{r TTEST1, echo=TRUE}
t.test(idata2.1$Mean.Temp)
```

T-test: sample 2

```{r TTEST2, echo=TRUE}
t.test(idata2.2$Mean.Temp)
```

T-test: sample 3

```{r TTEST3, echo=TRUE}
t.test(idata2.3$Mean.Temp)
```

