---
title: "Statistical Distributions"
author: "`r Sys.getenv('USERNAME')`"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
## basic settings
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE,
                      fig.height=4,
                      fig.width=8
                      )

## load common libraries
library(tidyverse)
library(gridExtra)
library(lubridate)
library(scales)
library(plotly)
library(PerformanceAnalytics)

## Reference: R Markdown cheatsheet
## https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf



```

## Distribution Exploration

Overview of major distributions and how to represent them with sample data.

[Good reference article](https://blog.cloudera.com/blog/2015/12/common-probability-distributions-the-data-scientists-crib-sheet/)

First, set some common distribution parameters that can be used across different examples:

### Parameters

```{r PARAM}
m <- 10 ## mean
s <- 2 ## std dev
nn <- 1000 ## number examples
pr <- 0.3 ## probability


set.seed(124)
```

*(not all distributions use all parameters and not all parameters are used by more than 1 distribution)*

### Normal 

```{r NORMAL}
## generate distribution sample
dname <- 'normal'
## set distribution parameters
dist <- rnorm(mean=m, sd=s, n=nn)

## set chart arrangement
par(mfrow=(c(1,2)))
## show data and histogram
plot(dist)
hist(dist)

## convert to data frame (done in hidden code chunks for subsequent distribution data)
df.dist <- data.frame(distribution=dname,
                      values=dist)
## set up data frame to collect all the distributions
df.dist.all <- df.dist
## create columnar data with distribution name
df.dist.col <- df.dist %>% select(values) %>% rename(
  !!dname:=values
)
## set up data frame to collect columnar data for all data frames
df.dist.all.cols <- df.dist.col

```



### Uniform

```{r UNI}
## generate distribution sample
dname <- 'uniform'

## set distribution parameters
dist <- runif(n=nn, min=m-s, max=m+s)
## set chart arrangement
par(mfrow=(c(1,2)))
## show data and histogram
plot(dist)
hist(dist)

```

```{r, echo=FALSE}
source('data-frame-process.R')
```

### Binomial

* based on `r nn` observations of 1 trial with probability `r pr`

```{r BINOM}
## generate distribution sample
dname <- 'binomial'

## set distribution parameters
## decrease nn because need only small number to show binomial
## and with more it stretches y-axis when comparing
dist <- rbinom(n=nn/10, size=1, prob = pr)
## set chart arrangement
par(mfrow=(c(1,2)))
## show data and histogram
plot(dist)
hist(dist)
```

```{r, echo=FALSE}
source('data-frame-process.R')
```

### Poisson

```{r POIS}
## generate distribution sample
dname <- 'poisson'
lam <- 10
## set distribution parameters
dist <- rpois(lambda=lam, n=nn)
## set chart arrangement
par(mfrow=(c(1,2)))
## show data and histogram
plot(dist)
hist(dist)

```

```{r, echo=FALSE}
source('data-frame-process.R')
```

### Summary

```{r SUMMARY CHARTS}

ggplot(df.dist.all, aes(x=values, color=distribution))+geom_density()
  
```

## Normality Testing

### Visually

```{r NORM1}

df.norm <- df.dist.all %>% filter(distribution=='normal') 

h <- ggplot(df.norm, aes(x=values))+geom_histogram(fill=NA, color='blue')

cd <- 
ggplot(df.norm, aes(x=values))+geom_density(color='red')

grid.arrange(h, cd, nrow=1)
```

...or better yet: overlay density of sample data with ideal normal curve for same mean and sd

```{r}
df.norm <- df.norm %>%
  mutate(normprob=dnorm(values, mean=m, sd=s))

ggplot(df.norm, aes(x=values))+geom_density(color='blue')+
  geom_line(aes(x=values, y=normprob), color='red')+theme_bw()

ggplot(df.norm, aes(x=values))+geom_histogram(aes(y=..density..), fill='lightblue', color='blue')+
  geom_line(aes(x=values, y=normprob), color='red')+theme_bw()
```

### Statistically

```{r NORM2}
## from: http://www.sthda.com/english/wiki/normality-test-in-r
shapiro.test(df.norm$values)
```

Null hypothesis (normality) is not refuted, as shown by high p-value.

### With other distributions

```{r}

df.poisson <- df.dist.all %>% mutate(normprob=dnorm(values, mean=lam, sd=s))

ggplot(df.poisson, aes(x=values))+geom_density(color='blue')+
  geom_line(aes(x=values, y=normprob), color='red')+theme_bw()

shapiro.test(df.poisson$values)

```

Comparing Poisson distributions with different levels of lambda

```{r}
df.pois.all <- data.frame()
for(i in 1:4){
  lam <- 10^i
  examplelam <- paste0('example',lam)
  pois <- rpois(n=100, lambda = lam)
  df.pois <- data.frame(example=examplelam,
                        values=pois)
  df.pois <- df.pois %>% mutate(
    normprob=dnorm(values, mean=lam, sd=sd(df.pois$values))
  )
  df.pois.all <- bind_rows(df.pois.all, df.pois)
}
## function for generating plots of all four variations
plotGen <- function(data, enum){
  data <- data %>% filter(example==paste0('example',enum))
  ggplot(data, aes(x=values))+geom_density(color='blue')+
    geom_line(aes(y=normprob), color='red', linetype='dashed')
} 
## use function to generate plots for each variation
pp100 <- plotGen(df.pois.all, 10)
pp200 <- plotGen(df.pois.all, 100)
pp300 <- plotGen(df.pois.all, 1000)
pp400 <- plotGen(df.pois.all, 10000)
## display plots in grid
grid.arrange(pp100, pp200, pp300, pp400, nrow=2)
```


