---
title: "Statistical Distributions"
author: "`r Sys.getenv('USERNAME')`"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
## basic settings
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE
                      )

## load common libraries
library(tidyverse)
library(gridExtra)
library(lubridate)
library(scales)
library(plotly)
library(PerformanceAnalytics)

## Reference: R Markdown cheatsheet
## https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf



```

## Distribution Exploration

Overview of major distributions and how to represent them with sample data.

[Good reference article](https://blog.cloudera.com/blog/2015/12/common-probability-distributions-the-data-scientists-crib-sheet/)

First, set some common distribution parameters that can be used across different examples:

### Parameters

```{r PARAM}
m <- 10 ## mean
s <- 2 ## std dev
nn <- 1000 ## number examples
pr <- 0.3 ## probability


set.seed(124)
```

*(not all distributions use all parameters and not all parameters are used by more than 1 distribution)*

### Normal 

```{r NORMAL}
## generate distribution sample
dname <- 'normal'
## set distribution parameters
dist <- rnorm(mean=m, sd=s, n=nn)

## set chart arrangement
par(mfrow=(c(1,2)))
## show data and histogram
plot(dist)
hist(dist)

## convert to data frame (done in hidden code chunks for subsequent distribution data)
df.dist <- data.frame(distribution=dname,
                      values=dist)
## set up data frame to collect all the distributions
df.dist.all <- df.dist
## create columnar data with distribution name
df.dist.col <- df.dist %>% select(values) %>% rename(
  !!dname:=values
)
## set up data frame to collect columnar data for all data frames
df.dist.all.cols <- df.dist.col

```



### Uniform

```{r UNI}
## generate distribution sample
dname <- 'uniform'

## set distribution parameters
dist <- runif(n=nn, min=m-s, max=m+s)
## set chart arrangement
par(mfrow=(c(1,2)))
## show data and histogram
plot(dist)
hist(dist)

```

```{r, echo=FALSE}
source('data-frame-process.R')
```

### Binomial

* based on `r nn` observations of 1 trial with probability `r pr`

```{r BINOM}
## generate distribution sample
dname <- 'binomial'

## set distribution parameters
dist <- rbinom(n=nn, size=1, prob = pr)
## set chart arrangement
par(mfrow=(c(1,2)))
## show data and histogram
plot(dist)
hist(dist)
```

```{r, echo=FALSE}
source('data-frame-process.R')
```

### Poisson

```{r POIS}
## generate distribution sample
dname <- 'poisson'

## set distribution parameters
dist <- rpois(lambda=10, n=1000)
## set chart arrangement
par(mfrow=(c(1,2)))
## show data and histogram
plot(dist)
hist(dist)

```

```{r, echo=FALSE}
source('data-frame-process.R')
```

### Summary

```{r SUMMARY CHARTS}

ggplot(df.dist.all, aes(x=values, color=distribution))+geom_density()
  

```

